---
title: "Coursera Data Science Capstone - Prediction model building"
author: "Samy Soualem"
date: "September 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initialization
```{r echo = FALSE, results = FALSE, message = FALSE}
library(quanteda)
library(ggplot2)
#library(doParallel)
library(data.table)
library(stringr)
library(foreach)
library(doSNOW)

source("model/global.R")
source("model/trainModel.R")

cores <- parallel::detectCores()

# Initialiaze list of DFM
lDfm <- vector(mode = "list", length = MAX_NGRAM_LEVEL)
dtDfmFull <- vector(mode = "list", length = MAX_NGRAM_LEVEL)

print(paste("Time :", Sys.time()))
```

## Files and directories
```{r echo = FALSE, results = FALSE}
# Data source : https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip
# Data directories
enUsDir <- "data/final/en_US"
enUsSampleDir <- file.path(enUsDir, paste0("sample_", FILE_SAMPLE_RATE))
enFileNames <- c("en_US.blogs.txt", "en_US.news.txt", "en_US.twitter.txt")
enFile <- file.path(enUsDir, enFileNames)


blogSampleFile <- file.path(enUsSampleDir
                  , paste("en_US.blogs_sample", FILE_SAMPLE_RATE, 1:NB_SUB_FILES, sep = "_")
                  )
newsSampleFile <- file.path(enUsSampleDir
                  , paste("en_US.news_sample", FILE_SAMPLE_RATE, 1:NB_SUB_FILES, sep = "_")
                  )
twitterSampleFile <- file.path(enUsSampleDir
                  , paste("en_US.twitter_sample", FILE_SAMPLE_RATE, 1:NB_SUB_FILES, sep = "_")
                  )

enSampleFile <- vector(mode = "list", length = NB_SUB_FILES)              

for(i in 1:NB_SUB_FILES) {
  enSampleFile[[i]] <- c(blogSampleFile[i], newsSampleFile[i], twitterSampleFile[i])
  enSampleFile[[i]] <- paste0(enSampleFile[[i]], ".txt")
}



rdsDir <- paste("data/rds", RDS_FILE_VERSION, paste("sample", FILE_SAMPLE_RATE, sep = '_') , sep = "/")
```

## RDS files preparation to save the results
```{r}
if (! dir.exists(rdsDir)) {
  dir.create(rdsDir, recursive = TRUE)
}

rdsSampleRate <- paste("sample", FILE_SAMPLE_RATE, sep ='_')
rdsSuffix <- ".rds"

# Version and sample rate in file name just in case
rdsFileSentences <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "sentenceTokens"
                                                   , 1:NB_SUB_FILES, sep = '_'), rdsSuffix))
rdsFileDtDfm <- vector(mode = "list", length = MAX_NGRAM_LEVEL)

rdsFileDtDfm[[1]] <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfm1", 1:NB_SUB_FILES, sep = '_'), rdsSuffix))
rdsFileDtDfm[[2]] <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfm2", 1:NB_SUB_FILES, sep = '_'), rdsSuffix))
rdsFileDtDfm[[3]] <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfm3", 1:NB_SUB_FILES, sep = '_'), rdsSuffix))
rdsFileDtDfm[[4]] <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfm4", 1:NB_SUB_FILES, sep = '_'), rdsSuffix))

rdsFileDtDfmFull <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfmFull", sep = '_'), rdsSuffix))
rdsFileDtDfmTrim <- file.path(rdsDir, paste0(paste(RDS_FILE_VERSION, rdsSampleRate, "dtDfmTrim", sep = '_'), rdsSuffix))
```

## Functions
### Function : load files as a Corpus
```{r}
# files = list of files to load as a Corpus
getCorpus <- function(files) {
   # Prepare parallel processing
  cl <- makeSOCKcluster(cores) # start a socket cluster for the numbers of cores available to enable those many parallel R processes
  registerDoSNOW(cl) # register the SNOW parallel backend with the foreach package
  
  # Load file fragments as a Corpus list
  print("Loading file fragments as a Corpus")
  pb <- txtProgressBar(max=NB_SUB_FILES, style = 3)
  progress <- function(n) setTxtProgressBar(pb, n)
  opts <- list(progress=progress)
  corpusList <- foreach(i = 1:NB_SUB_FILES,
                        .packages=c("quanteda"), # mention all packages used in the code block
                        .options.snow=opts) %dopar% {
                    
    corpus(textfile(files[[i]], encoding = "UTF-8"))
  }    
  close(pb)
  # Stop all clusters
  stopCluster(cl)
  
  corpusList
}
```

### Function : tokenize sentences in parallel
```{r}
# corpus : list of corpus
getSentenceTokens <- function(corpus) {
  # Prepare parallel processing
  cl <- makeSOCKcluster(cores) # start a socket cluster for the numbers of cores available to enable those many parallel R processes
  registerDoSNOW(cl) # register the SNOW parallel backend with the foreach package
  
  # Tokenize sentences
  print("Tokenizing sentences")
  pb <- txtProgressBar(max=NB_SUB_FILES, style = 3)
  progress <- function(n) setTxtProgressBar(pb, n)
  opts <- list(progress=progress)
  sentenceTokens <- foreach(i = 1:NB_SUB_FILES,
                        .packages=c("quanteda"), # mention all packages used in the code block
                        .export = "tokenizeSentences",
                        .options.snow=opts) %dopar% {
                    
    tokenizeSentences(corpus[[i]])
  }    
  close(pb)
  
  # Stop all clusters
  stopCluster(cl)
  sentenceTokens
}
```

### Function : generate n-grams in parallel
```{r}
# tokens : sentence tokens
# n : n-gram level
generateNgramFiles <- function(tokens, n) {
  # Prepare parallel processing
  cl <- makeSOCKcluster(cores) # start a socket cluster for the numbers of cores available to enable those many parallel R processes
  registerDoSNOW(cl) # register the SNOW parallel backend with the foreach package
  
  print(paste0("Tokenizing ", n,"-grams"))
  pb <- txtProgressBar(max=length(tokens), style = 3)
  progress <- function(n) setTxtProgressBar(pb, n)
  opts <- list(progress=progress)
  null <- foreach(i = 1:length(tokens),
                        .packages=c("quanteda", "data.table"), # mention all packages used in the code block
                        .export  = c("rdsFileDtDfm", "generateNgramDt", "generateDfm", 
                                     "splitNgramPrefix", "getSwearWords", "SENT_START_MARKER"),
                        .options.snow=opts) %dopar% {
    
    if(n == 1) {
      ignoredFeatures <- c(getSwearWords(), SENT_START_MARKER)
    } else {
      ignoredFeatures <- getSwearWords()
    }
    
    dtDfm <-generateNgramDt(tokens[[i]], n, ignoredFeatures)
    saveRDS(dtDfm, rdsFileDtDfm[[n]][[i]])
    rm(dtDfm)
    gc()
  }    
  close(pb)
  
  # Stop all clusters
  stopCluster(cl)
}
```

### Function : generate n-grams sequentially
```{r}
# tokens : sentence tokens
# n : n-gram level
generateNgramFilesSequential <- function(tokens, n) {
  pb <- txtProgressBar(max=NB_SUB_FILES, style = 3)
  
  for(i in 1:NB_SUB_FILES) {
    print(paste0("Tokenizing ", n,"-grams for sample number ", i))
    if(n == 1) {
      ignoredFeatures <- c(getSwearWords(), SENT_START_MARKER)
    } else {
      ignoredFeatures <- getSwearWords()
    }
    
    dtDfm <-generateNgramDt(tokens[[i]], n, ignoredFeatures)
    saveRDS(dtDfm, rdsFileDtDfm[[n]][[i]])
    rm(dtDfm)
    gc()
    setTxtProgressBar(pb, i)
  }    
  close(pb)
}
```


### Function : load DFM fragments from RDS files
```{r}
# n = n-gram level
# files = vector of RDS files to load
getDtDfmFragments <- function(n, rdsFiles) {
  # Prepare parallel processing
cl <- makeSOCKcluster(cores) # start a socket cluster for the numbers of cores available to enable those many parallel R processes
registerDoSNOW(cl) # register the SNOW parallel backend with the foreach package

# Read n-grams data tables files
print(paste("Loading files :", rdsFiles))
pb <- txtProgressBar(max=NB_SUB_FILES, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress=progress)
dfmFragments <- foreach(i = 1:NB_SUB_FILES,
                      .packages=c("data.table"), # mention all packages used in the code block
                      .options.snow=opts) %dopar% {
  
  # Load n-grams
  #readRDS(rdsFileDtDfm[[n]][i])
  readRDS(rdsFiles[i])
}

close(pb)
gc()

# Stop all clusters
stopCluster(cl)

dfmFragments
}
```

## Creating sampled files

Total file sample rate : `r I(FILE_SAMPLE_RATE * 100)`%.


```{r echo = FALSE, results = FALSE}
print(paste("Time :", Sys.time()))

# Stop with an error if the original files are not found
if (! all(file.exists(enFile))) {
  errMsg <- paste("The following files were not found :", paste(enFile, collapse = ", "))
  stop(errMsg)
}


# If all the sample files don't already exist for this sample rate, create them
if (! all(file.exists(unlist(enSampleFile)))) {
  if (! dir.exists(enUsSampleDir)) {
    dir.create(enUsSampleDir)
  }
  
  # Read all files, generate a random sample and save it to another file for later use
  
  
  set.seed(1234)
  for(i in 1:length(enFile)) {
    print(paste("Loading file :", enFile[i]))
    
    # Read whole file
    fullfile <- readLines(enFile[i])
    
    # Split into NB_SUB_FILES fragments
    totalSampleNbLines <- ceiling(SUB_FILE_SAMPLE_RATE * NB_SUB_FILES * length(fullfile))
    tempData <- sample(fullfile, totalSampleNbLines)
    
    subFileNbLines <- totalSampleNbLines / NB_SUB_FILES
    
    splitFactor <- ceiling(seq_along(tempData) / subFileNbLines)
    subFileData <- split(tempData, splitFactor)
    
    # If there is a chunk with the number (NB_SUB_FILES + 1), add it to the NB_SUB_FILES element
    # to have exactly NB_SUB_FILES files
    if(length(subFileData) > NB_SUB_FILES) {
      subFileData[[NB_SUB_FILES]] <- c(subFileData[[NB_SUB_FILES]], subFileData[[NB_SUB_FILES + 1]])
    }
    
    print(paste("splitting and saving file :", enFile[i], "into", NB_SUB_FILES, "fragments"))
    pb <- txtProgressBar(max=NB_SUB_FILES, style = 3)
    # Save all sub file samples to files
    for(j in 1:NB_SUB_FILES) {
      writeLines(subFileData[[j]], enSampleFile[[j]][i])
      setTxtProgressBar(pb, j)
    }
    close(pb)
  }
  # Free memory (not very useful in general apparently)
  rm(fullfile)
  rm(tempData)
  rm(subFileData)
  gc()  # garbage collector
}
# Sample files here, nothing to do

print(paste("Time :", Sys.time()))
```





## Generate DFM fragments and save them to RDS files
```{r}
# If the DFM fragments files are not there, generate and save them to files
if (! all(file.exists(unlist(rdsFileDtDfm)))) {
  print(paste("Time :", Sys.time()))
  enRawCorpus <- getCorpus(enSampleFile)
  print(paste("Time :", Sys.time()))
  
  sentenceTokens <- getSentenceTokens(enRawCorpus)
  print(paste("Time :", Sys.time()))
  rm(enRawCorpus)
  gc()
  
  # Limit number of parallel tasks for n-gram creation because of RAM limitations
  maxParallelNb <- 2
  nbIterations <- ceiling(NB_SUB_FILES / maxParallelNb)
  
  # Generate DFM for each n-gram level and save them to RDS files
  for(i in 1:MAX_NGRAM_LEVEL) {
    print(paste("Time :", Sys.time()))
    
    # Another loop level to avoid processing everything in parallel at once
    for(j in 0:(nbIterations - 1)) {
      print(paste("Creating n-grams with n =", i, "- Iteration", j + 1, "out of", nbIterations))
      
      idxStart <- maxParallelNb * j + 1
      idxEnd <- maxParallelNb * j + maxParallelNb
      
      if(idxEnd > NB_SUB_FILES) {
        idxEnd <- NB_SUB_FILES
      }
      
      generateNgramFiles(sentenceTokens[idxStart:idxEnd], i)
      gc()
    }
    
    # Backup method : do everything in sequence 
    #generateNgramFilesSequential(sentenceTokens, i)
    gc()
  }
  rm(sentenceTokens)
  gc()
  
} else {
  # do nothing
}
print(paste("Time :", Sys.time()))
```





## Aggregate smaller 1-grams data tables
```{r}
if (!file.exists(rdsFileDtDfmFull)) {

  for(n in 1:MAX_NGRAM_LEVEL) {
    # Load DFM fragments files
    print(paste("Time :", Sys.time()))
    tempDfm <- getDtDfmFragments(1, rdsFileDtDfm[[n]])
    print(paste("Time :", Sys.time()))
  
    # Init with first data.table
    dtDfmFull[[n]] <- tempDfm[[1]]
    
    # Aggregate smaller n-grams data tables
    print(paste0("Merging ", n, "-grams DFM"))
    colToRm <- c("count.x", "count.y")
    for(i in 2:NB_SUB_FILES) {
      # Merge the data tables by ngram (prefix + lastWord) => full outer join
      dtDfmFull[[n]] <- merge(dtDfmFull[[n]], tempDfm[[i]], by = c("prefix", "lastWord"), all = TRUE)
      
      # transform NA into 0
      dtDfmFull[[n]][is.na(count.x), count.x := 0]
      dtDfmFull[[n]][is.na(count.y), count.y := 0] 
      
      # Add the ngram counts
      dtDfmFull[[n]][, count := count.x + count.y] 
      
      # Remove individual count results
      dtDfmFull[[n]][, (colToRm) := NULL] 
    }
  }
  print(paste("Saving merged full DFM to file", rdsFileDtDfmFull))
  saveRDS(dtDfmFull, rdsFileDtDfmFull)
  gc()
} else {
  dtDfmFull <- readRDS(rdsFileDtDfmFull)
}

print(paste("Time :", Sys.time()))
```




## Remove low-frequency n-grams to reduce model size and lower response time
```{r}
if (!file.exists(rdsFileDtDfmTrim)) {
  dtDfmTrim <- vector(mode = "list", length = MAX_NGRAM_LEVEL)
  
  print(paste("Trimming DFM. MIN_NGRAM_COUNT =", MIN_NGRAM_COUNT))
  # Delete low frequency ngram
  for(i in 1:MAX_NGRAM_LEVEL) {
    dtDfmTrim[[i]] <- dtDfmFull[[i]][count >= MIN_NGRAM_COUNT, ]
  }
  print(paste("Saving trimmed DFM to file", rdsFileDtDfmTrim))
  saveRDS(dtDfmTrim, rdsFileDtDfmTrim)
} else {
  dtDfmTrim <- readRDS(rdsFileDtDfmTrim)
}
print(paste("Time :", Sys.time()))
```
